---
title: "Calculating Site Risk Scores"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating Site Risk Scores}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

This vignette describes how **site-level risk scores** are calculated using Key Risk Indicators (KRIs), their associated *flag severities*, and predefined *weights*. The risk scoring framework provides a standardized, quantitative way to highlight sites with elevated risk.

# Risk Indicators and Flags

Each KRI (e.g., Adverse Events, Serious Adverse Events, Protocol Deviations) can generate *flags* of different severity:
  
  * **Low Amber**
  * **Low Red**
  * **High Amber**
  * **High Red**
  
  Each flag type is assigned a **weight** that reflects its contribution to site risk.

```{r, echo=FALSE}
library(knitr)

kri_weights <- data.frame(
  KRI = c("AE", "SAE", "PD (Protocol Deviations)", "IPD (Important Protocol Deviations)",
          "LB (Labs)", "Data Quality (combined)", "Screen Failure",
          "Treatment Discontinuation", "Study Discontinuation"),
  Low_Amber = c(2, 0, 0.5, 0, 0, 0, 0, 0, 0),
  Low_Red   = c(4, 1, 1, 0, 0, 0, 0, 0, 0),
  High_Amber= c(0, 0.5, 1, 2, 0, 0.5, 1, 2, 2),
  High_Red  = c(0, 1, 2, 4, 0, 1, 2, 4, 4)
)

kable(kri_weights, caption = "Flag Weights for Key Risk Indicators")
```

# Example Site Calculation

For each site, the observed *flags* are multiplied by the corresponding **weights** and summed to produce a **site risk score**.

Below is a worked example for one site:
  
```{r, echo=FALSE}
site_example <- data.frame(
  KRI = c("AE", "SAE", "PD", "IPD", "LB", "Data Quality", 
          "Screen Failure", "Treatment Discontinuation", "Study Discontinuation"),
  Contribution = c(4.0, 0.0, 1.0, 2.0, 0.0, 0.5, 2.0, 0.0, 0.0)
)

kable(site_example, caption = "Example Site Contributions to Risk Score")
```

The **total site risk score** is simply the sum:
  
```{r}
total_score <- sum(site_example$Contribution)
total_score
```

# Benchmarking Across Sites

Site-level scores can be contextualized in two ways:
  
* **Relative ranking**: comparing each site’s score against other sites (e.g., percentile, rank order).
* **Temporal averaging**: computing average scores across multiple months to reduce short-term fluctuations.

For example:
  
  * Highest site score observed: **22**
  * Average score across 3 months: **9.5**
  
  # Conclusion
  
This scoring framework provides an interpretable, quantitative measure of risk at each site. By combining information across KRIs and weighting severity appropriately, it allows data monitoring teams to identify high-risk sites and prioritize monitoring resources effectively.

