---
title: "Cookbook - Reporting Workflow Examples"
type: "Cookbook"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    self_contained: true
    theme: cosmo
---

```{r rmd-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,
  message = TRUE,
  warning = TRUE,
  results = 'markup',
  comment = "#>"
)
```

```{r load-packages}
library(dplyr)
library(gsm.core)
library(gsm.mapping)
library(gsm.reporting)
library(gsm.kri)
library(yaml)
```

## Example 3.1 - Create a KRI Report using 12 standard metrics in a step-by-step workflow

This example demonstrates a complete workflow from raw data to KRI reports using standard metrics.

### Prepare Raw Data

```{r prepare-raw-data}
core_mappings <- c(
  "AE", "COUNTRY", "DATACHG", "DATAENT", "ENROLL", "EXCLUSION", "IE", "LB", "VISIT",
  "PD", "PK", "QUERY", "STUDY", "STUDCOMP", "SDRGCOMP", "SITE", "SUBJ"
)

lRaw <- list(
  Raw_SUBJ = gsm.core::lSource$Raw_SUBJ,
  Raw_AE = gsm.core::lSource$Raw_AE,
  Raw_PD = gsm.core::lSource$Raw_PD %>%
    rename(
      subjid = subjectenrollmentnumber,
      dvdecod = crocategory,
      dvterm = description,
      dvdtm = deviationdate
    ),
  Raw_LB = gsm.core::lSource$Raw_LB,
  Raw_PK = gsm.core::lSource$Raw_PK %>%
    rename(
      visit = foldername
    ),
  Raw_STUDCOMP = gsm.core::lSource$Raw_STUDCOMP,
  Raw_SDRGCOMP = gsm.core::lSource$Raw_SDRGCOMP,
  Raw_DATACHG = gsm.core::lSource$Raw_DATACHG %>%
    rename(subject_nsv = subjectname),
  Raw_DATAENT = gsm.core::lSource$Raw_DATAENT %>%
    rename(subject_nsv = subjectname),
  Raw_QUERY = gsm.core::lSource$Raw_QUERY %>%
    rename(subject_nsv = subjectname),
  Raw_ENROLL = gsm.core::lSource$Raw_ENROLL,
  Raw_SITE = gsm.core::lSource$Raw_SITE %>%
    rename(
      studyid = protocol,
      invid = pi_number,
      InvestigatorFirstName = pi_first_name,
      InvestigatorLastName = pi_last_name,
      City = city,
      State = state,
      Country = country
    ),
  Raw_STUDY = gsm.core::lSource$Raw_STUDY %>%
    rename(studyid = protocol_number),
  Raw_VISIT = gsm.core::lSource$Raw_VISIT %>%
    rename(visit = foldername),
  Raw_IE = gsm.core::lSource$Raw_IE
)
```

### Step 1 - Create Mapped Data Layer

Filter, aggregate and join raw data to create mapped data layer.

```{r create-mapped-data, results='hide', warning=FALSE, message=FALSE}
mappings_wf <- MakeWorkflowList(strNames = core_mappings, strPath = "workflow/1_mappings", strPackage = "gsm.mapping")
mapped <- RunWorkflows(mappings_wf, lRaw)
```

### Step 2 - Create Metrics

Calculate metrics using mapped data.

```{r create-metrics, results='hide', warning=FALSE, message=FALSE}
metrics_wf <- MakeWorkflowList(strPath = "inst/workflow/2_metrics", strPackage = "gsm.kri")
analyzed <- RunWorkflows(metrics_wf, c(mapped, list(lWorkflows = metrics_wf)))
```

### Step 3 - Create Reporting Layer

Create reports using metrics data.

```{r create-reporting, results='hide', warning=FALSE, message=FALSE}
reporting_wf <- MakeWorkflowList(strPath = "workflow/3_reporting", strPackage = "gsm.reporting")
reporting <- RunWorkflows(reporting_wf, c(mapped, list(
  lAnalyzed = analyzed,
  lWorkflows = metrics_wf
)))
```

### Step 4 - Create KRI Reports

Create KRI Site and Country reports using reporting data.

```{r create-kri-reports, results='hide', warning=FALSE, message=FALSE}
module_wf <- MakeWorkflowList("kri", strPath = "inst/workflow/4_modules", strPackage = "gsm.kri")
lReports <- RunWorkflows(module_wf, reporting)
```

## Example 3.3 - Site-Level KRI Report with multiple SnapshotDate

This example demonstrates creating visualizations and reports using pre-calculated reporting data from `gsm.core`.

### Creating Interactive Charts

```{r create-charts-3-3}
lCharts <- MakeCharts(
  dfResults = gsm.core::reportingResults,
  dfGroups = gsm.core::reportingGroups,
  dfMetrics = gsm.core::reportingMetrics,
  dfBounds = gsm.core::reportingBounds
)
```

### Generate KRI Report

```{r generate-report-3-3, results='hide', warning=FALSE, message=FALSE}
kri_report_path <- Report_KRI(
  lCharts = lCharts,
  dfResults = FilterByLatestSnapshotDate(gsm.core::reportingResults),
  dfGroups = gsm.core::reportingGroups,
  dfMetrics = gsm.core::reportingMetrics
)
```


## Example 3.4 - Reporting Results with Changes from previous snapshot

This example shows how to include historical comparison data to highlight changes from previous snapshots.

### Prepare Historical Data

```{r historical-data-3-4}
historical <- gsm.core::reportingResults %>% filter(SnapshotDate == "2025-03-01")
```

### Re-run reporting model and KRI report with historical data

```{r reporting-with-historical-3-4, results='hide', warning=FALSE, message=FALSE}
reporting_long <- gsm.core::RunWorkflows(reporting_wf, c(mapped, list(lAnalyzed = analyzed, Reporting_Results_Longitudinal = historical, lWorkflows = metrics_wf)))
lReports_long <- gsm.core::RunWorkflows(module_wf, reporting_long)
```

